BayesianAnalysisForPrereg

### 2018-04-11 E2/analysis/bayesianAnalysisForPrereg.Rmd
good

altHypothesisPrior=function(delta) {
  mean= meanCohensD
  sigma= sd(cohensDs)
  y= dnorm(delta,mean,sigma)
  y=y*as.integer(delta>lo)*as.integer(delta<hi) #multiply by zero if outside of support
}
#Normalize alternative density in case is not already
K=1/integrate(altHypothesisPrior,lower=lo,upper=hi)$value
f=function(delta) K*altHypothesisPrior(delta)

@@

#Plot Alternative as a density and Null as a point arrow
maxAlt=max(f(delta))
plot(dDomain,f(dDomain),typ='n',xlab="Left/top bias difference (Cohen's d)",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
lines(dDomain,f(dDomain),col='green',lwd=2)
legend("topright",legend=c("Null","Alternative","Previous findings"),col=c('darkblue','green','black'),lwd=2)

@@

altPredIntegrand=function(delta,obs,N) {
  #call the student-t distribution, with delta the effect size 
  dt(sqrt(N)*obs,N-1,sqrt(N)*delta)*f(delta)
}

@@

```{r}  
smaller=.3333
my.es= meanCohensD * smaller
my.sd = sd(cohensDs)
```

Assume effect size is only `r smaller` as big as original `r meanCohensD`, yielding `r smaller`
 . So, both set prior to half as big and show likelihood functions for whole range but highlight that effect size, for a few different sample sizes.
For standard deviation of prior  will use `r my.sd`
```{r alternativePrior}

@@

N=75
```
Have set alternative prior as smaller effect.

@@

altPredF=function(obsEffectSizes,N) {
  suppressWarnings( integrate(altPredIntegrand,lower=lo,upper=hi,obs=obsEffectSizes,N=N)$value
  )
}

@@

plot(type='l',obsEffectSizes,nullPred,ylim=c(0,top),xlab="Observed Effect Size",ylab="Density",main="P(data|H)",col='darkblue',lwd=2)
lines(obsEffectSizes,altPred,col='darkgreen',lwd=2)
legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
text(-.9,.35,paste0("n=",N))
#plot example small effect size points on curve
abline(v=my.es,lty=2,lwd=2,col='red')
valNull=nullPredF(my.es,N)
valAlt=altPredF(my.es,N)
points(pch=19,c(my.es,my.es),c(valNull,valAlt),col="red")
text(my.es+.1,valAlt+.1,paste0("BF=",round(valAlt/valNull,2)),col="red")

#plot example large effect size points on curve
large.es<-1
abline(v=large.es,lty=2,lwd=2,col='red')
valNull=nullPredF(large.es,N)
valAlt=altPredF(large.es,N)
points(pch=19,c(large.es,large.es),c(valNull,valAlt),col="red")
text(large.es+.3,valAlt+.1,paste0("BF=",round(valAlt/valNull,2)),col="red")
```
And for n=100
```{r subjexts100}
#For each effect size, compute density
obsEffectSizes=seq(-1,1.5,.01)

I=length(obsEffectSizes)
altPred=1:I
N=100
for (i in 1:I) {
  altPred[i] = altPredF(obsEffectSizes[i],N)
}

@@

#plot example effect size points on curve
abline(v=my.es,lty=2,lwd=2,col='red')
valNull=nullPredF(my.es,N)
valAlt=altPredF(my.es,N)
points(pch=19,c(my.es,my.es),c(valNull,valAlt),col="red")
text(my.es+.2,valAlt+.1,paste0("BF=",round(valAlt/valNull,2)),col="red")
```

@@

Do the correlogram and correlation

>>>>

altHypothesisPrior=function(delta) {
  mean= meanCohensD
  sigma= sd(cohensDs)
  y= dnorm(delta,mean,sigma)
  y=y*as.integer(delta>lo)*as.integer(delta<hi) #multiply by zero if outside of support
}
#Normalize alternative density in case is not already
K=1/integrate(altHypothesisPrior,lower=lo,upper=hi)$value
prior=function(delta) K*altHypothesisPrior(delta)

@@

#Plot Alternative as a density and Null as a point arrow
maxAlt=max(prior(delta))
plot(dDomain,prior(dDomain),typ='n',xlab="Left/top bias difference (Cohen's d)",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
lines(dDomain,prior(dDomain),col='green',lwd=2)
legend("topright",legend=c("Null","Alternative","Previous findings"),col=c('darkblue','green','black'),lwd=2)

@@

altPredIntegrand=function(delta,obs,N) {
  #call the student-t distribution, with delta the effect size 
  dt(sqrt(N)*obs,N-1,sqrt(N)*delta)*prior(delta)
}

@@

```{r}  
smaller=.3333
my.es= meanCohensD * smaller
my.sd = sd(cohensDs) * 1.5
```

Assume effect size is only `r smaller` as big as original `r meanCohensD`, yielding `r my.es`
 . So, both set prior to half as big and show likelihood functions for whole range but highlight that effect size, for a few different sample sizes.
For standard deviation of prior  will use `r my.sd`
```{r alternativePrior}

@@

#Plot this new smaller-effect prior
dDomain=seq(-.3,2,.01)

#Plot Alternative as a density and Null as a point arrow
maxAlt=max(altPriorNormalized(delta))
plot(dDomain,altPriorNormalized(dDomain),typ='n',xlab="Left/top bias difference (Cohen's d)",ylab="Density",ylim=c(0,1.4*maxAlt),main="New prior putting more density over smaller effect sizes")
arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
lines(dDomain,altPriorNormalized(dDomain),col='green',lwd=2)
legend("topright",legend=c("Null","Alternative","Previous findings"),col=c('darkblue','green','black'),lwd=2)

points(cohensDs, rep(.1,4),pch=19) #plot previous effect sizes
text(cohensDs, c(.25,.35,.25,.25), labels,cex=.75)
```

``` {r}
N=75
```

@@

altPredF=function(obs,N) {
  suppressWarnings( integrate(altPredIntegrand,lower=lo,upper=hi,obs=obs,N=N)$value
  )
}

@@

plot(type='l',obsEffectSizes,nullPred,ylim=c(0,top),xlab="Observed Effect Size",ylab="Density",main="P(data|H)",col='darkblue',lwd=2)
lines(obsEffectSizes,altPred,col='darkgreen',lwd=2)
legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
text(-.9,.35,paste0("n=",N))
#plot example small effect size point on curve
abline(v=my.es,lty=2,lwd=2,col='red')
valNull=nullPredF(my.es,N)
valAlt=altPredF(my.es,N)
points(pch=19,c(my.es,my.es),c(valNull,valAlt),col="red")
text(my.es+.1,valAlt+.1,paste0("BF=",round(valAlt/valNull,1)),col="red")

#plot example even smaller effect size points on curve
abline(v=my.es-.1,lty=2,lwd=2,col='red')
valNull=nullPredF(my.es-.1,N)
valAlt=altPredF(my.es-.1,N)
points(pch=19,c(my.es-.1,my.es-.1),c(valNull,valAlt),col="red")
text(my.es-.2,valAlt+.1,paste0("BF=",round(valAlt/valNull,1)),col="red")

#plot example large effect size points on curve
large.es<-1
abline(v=large.es,lty=2,lwd=2,col='red')
valNull=nullPredF(large.es,N)
valAlt=altPredF(large.es,N)
points(pch=19,c(large.es,large.es),c(valNull,valAlt),col="red")
text(large.es+.3,valAlt-.1,paste0("BF=",round(valAlt/valNull,1)),col="red")
```

```{r echo=FALSE, message=FALSE}
N=100
```

And for n=`r N`

```{r subjexts100, echo=FALSE}
#For each effect size, compute density
obsEffectSizes=seq(-1,1.5,.01)

@@

#plot example effect size points on curve
abline(v=my.es,lty=2,lwd=2,col='red')
valNull=nullPredF(my.es,N)
valAlt=altPredF(my.es,N)
points(pch=19,c(my.es,my.es),c(valNull,valAlt),col="red")
text(my.es+.2,valAlt+.1,paste0("BF=",round(valAlt/valNull,1)),col="red")

#plot example even smaller effect size points on curve
abline(v=my.es-.1,lty=2,lwd=2,col='red')
valNull=nullPredF(my.es-.1,N)
valAlt=altPredF(my.es-.1,N)
points(pch=19,c(my.es-.1,my.es-.1),c(valNull,valAlt),col="red")
text(my.es-.15,valAlt,paste0("BF=",round(valAlt/valNull,1)),col="red")

So in summary, the above two graphs with associated blocks of code show how the Bayes factor will be calculated for whatever observed effect size turns up, and is thus suitable for citing in the prereg document. To boil it down, the prior we'll be using is simply a Gaussian truncated at zero (bright green curve several plots ago with mean Cohen's d = `r my.es` and sigma = `r my.sd` but it does mean that even with n=100, an observed effect size of `r round(my.es-.1,2)` means the evidence is very ambiguous (Bayes factor close to 1, specifically `r round(valAlt/valNull,1)` )

@@
